---
title: "RWorksheet_Huervana#4B"
output: pdf_document
date: "2023-10-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1. 

It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.

```{r name, echo = TRUE}


vectorA <- c(1,2,3,4,5)

matrx <- matrix(0,nrow = 5, ncol =5)

for (j in 1:5)
  for (k in 1:5)
  {
    matrx[j,k] <- abs (vectorA[j] - vectorA[k])
  }

matrx


```

2. Print the string "*" using for() function. The output should be the same as shown
in Figure 2

```{r name, echo = TRUE}

for(i in 1:5) {
  stars <- rep("*", i)
  print(stars)
}

```

3. Get an input from the user to print the Fibonacci sequence starting from the 1st input
up to 500. Use repeat and break statements. Write the R Scripts and its output.

```{r name, echo = TRUE}

input <- as.integer(readline("Enter number sequence starting from the 1st input up to 500:  "))

x <- input
y <- 0


cat("Fibonacci sequence starting from", input, ":\n")

repeat {
  
  nextSeq <- x + y
  
  if (nextSeq > 500){
    break
  }
  cat(nextSeq, " ")
  x <- y
  y <- nextSeq
  
}

```

4. Import the dataset as shown in Figure 1 you have created previously.

4a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? 
Show your codes and its result.

```{r name, echo = TRUE}

write.csv(HouseholdData, file = "HouseholdData.csv")

importData<- read.csv("HouseholdData.csv")
importData
import<-head(HouseholdData, 6)
import

```













